
## About myself

<!--- setup r lib -->
```{r}
#| label: "setup"
#| echo: false
#| include: false
#| eval: true

#library(knitr)
library(cowplot)
library(ExperimentHub)
```


- PhD student in Bioinformatics under Dr. Amrit Singh supervision

- BSc. in Math + minor in Data Science

- What does "bioinformatician" do? 
  - Assist researchers like you to better understand what you're data means
  - ~~We're just coders that know little bit more bio~~ ...
  
- Currently working in creating pipeline, tools, models analyzing biological data in an automated-fashion
  - Focused on machine learning & [AI]{.secondary}
  - [Reproducible]{.tertiary} workflows 

  


## Single Cell RNA-seq

What is single cell RNA sequencing?





## Loading data

The data is from a study [@kang2018multiplexed] and publicly avaliable  through R's `ExperimentHub` function


```{r}
#| label: "load-data"
#| echo: true
#| eval: true


eh <- ExperimentHub() #  Initialize the hub as some list object
sce <- eh[["EH2259"]] # We could then extract the match entry by taking this entry from out hub
# Then print it
sce
```

- After loaded data should inspect basic information
- What are rows? column? size of data?
- [Rows]{.tertiary} = genes
- [Columns]{.secondary} = cells

## Preprocessing of data before analysis

The data retrieved is [rawest form]{.secondary}, not all of it is useful.

Caveats:

::: {.incremental}

- Undetected genes
- Cells with very few or many detected genes
- Lowly expressed genes
- unnormalized expression values

:::




::: {.fragment}

This is usually the **quality control** (QC) step. This could potentially be another tutorial, so not deeply covered today.

We only perfom simple actions

:::

::: {.notes}

- QC to remove damaged cells, poor sequencing libraries,  error

:::

## Remove undetected genes

```{r}
#| label: "remove-undetected"
#| echo: true
#| eval: true

counts(sce) |> colnames() |> sample(10)

```




## Reference

